Redundant calibration as a complex optimization problem

### 2017-05-04 paper.tex
started implementing Gianni's comments

\section{Wirtinger Calibration}
\label{sec:p_statement}
%In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
%equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
%We present the signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.

%\subsection{Scalar Measurement Equation}
%\label{sec:s_mes}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Eq.~\eqref{eq:vis_definition} is known as the scalar measurement equation. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We normally try to estimate the true value of
$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.

@@

Eq.~\eqref{eq:vis_definition} can also be expressed in the following vector form 
\begin{equation}
\label{eq:vis_definition}
\bd = \bv + \bn, 
\end{equation}
where 
\begin{align}
 \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{pq} \conj{g_q},\nonumber\\
 \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
\end{align}
The function $\alpha_{pq}$ maps composite antenna indexes to unique single indexes and is defined as
\begin{equation}
\alpha_{pq} =
\begin{cases}
(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
0 & \textrm{otherwise}
\end{cases}.
\end{equation}
For example, the set
\begin{equation}
\{\alpha_{12},\alpha_{13},\cdots,\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\}
\end{equation}
and the set
\begin{equation}
\{1,2,\cdots,N-1,N,\cdots,B\}
\end{equation}
are equal. We denote the number of baselines in an array with $B$, i.e. $B = \frac{N^2-N}{2}$. The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).

We can now formulate calibration as an optimization problem:
\begin{equation}
\label{eq:least_squares}
\min_{\bg} \Lambda(\bg) = \min_{\bg} \|\br\|_F^2 = \min_{\bg} \|\bd - \bv(\bg)\|_F^2, 
\end{equation}
where $\bd$ is the observed visibility vector, $\bv$ is the model predicted visibility vector and $\bg =[g_1,\cdots,g_N]^T$. Closely inspecting Eq.~\eqref{eq:least_squares},
we realize that it is in actual fact a least-squares problem, i.e. $\Lambda$ is the objective function, $\br$ is the residual vector, $\bd$ is the observed data vector, $\bv$ is the 
model-predicted data vector and $\bg$ is the model-parameter vector that minimizes the objective function.

The gradient-based minimization algorithms that are generally used to solve least squares problems generally require the model (in Eq.~\eqref{eq:least_squares} the model is denoted by $\bv$) to be differentiable
towards each model parameter, i.e. these methods make use of a Jacobian matrix $\bJ$. In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
When the least squares problem is complex it becomes less straightforward to apply these gradient-based minimization methods, as
many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.

The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.

Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
are defined to be 
\begin{align}
\label{eq:wir}
\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ). 
\end{align}
Using the above definitions we can now easily compute the following:
\begin{align}
\frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
\end{align}
In this alternative approach we treat the complex parameters and their conjugates as separate variables.

\citet{Smirnov2015} recently proposed to use the Wirtinger derivatives to perform normal skymodel-based calibration. Eq.~\eqref{eq:least_squares} can be recast to
\begin{equation}
\label{eq:least_squares_augmented}
\min_{\breve{\bg}} \Lambda(\breve{\bg}) = \min_{\breve{\bg}} \|\breve{\br}\|_F^2 = \min_{\breve{\bg}} \|\breve{\bd} - \breve{\bv}(\breve{\bg})\|_F^2, 
\end{equation}
where $\breve{\br} = [\br^T,\conj{\br}^T]^T$, $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$. 
The complex GN update step is:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bg} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br},
\end{equation}
with 
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
and
\begin{align}
[\bmJ]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial g_i}, & [\bmJ^*]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial \conj{g}_i}. 
\end{align}
We refer to $\bJ$ as the Jacobian matrix. Note that the Wirtinger derivatives (see Eq.~\eqref{eq:wir}) are employed in Eq.~\eqref{eq:Jacobian}.
The complex LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bg} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}
where $\bD=\bI\odot\bJ^H\bJ$. We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$.

Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} is used to iteratively update the 
parameter vector $\breve{\bg}$:
\begin{equation}
\label{eq:update}
\breve{\bg}_{k+1} = \breve{\bg}_{k} + \Delta \breve{\bg}_{k}. 
\end{equation}
The parameter vector is updated until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
Eq.~\eqref{eq:GN_update} and Eq.~\eqref{eq:LM_update}.

@@

@@

\subsection{Signal-to-noise Ratio}
\label{sec:snr}
The following SNR definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

\subsection{Limitations}
\label{sec:scope}
We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}. Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.

\section{Redundant Calibration}
\label{sec:main_sec}
As discussed in Section~\ref{sec:ls}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
imaginary parts and then solve for each part of the parameters seperately or we can employ the notion of Wirtinger derivatives. The former 
approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.

\subsection{Real and imaginary Framework}
\label{sec:ri}
As we mentioned in Section~\ref{sec:ls} using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
splitting the problem into its phase-and-amplitude components.

@@

\subsection{Complex Framework}
\label{sec:w}

\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
that the Wirtinger derivative can also be applied to redundant calibration.

%We can perform redundant calibration by solving the following optimization problem: 
%\begin{equation}
%\label{eq:least_squares_red}
%\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
%\end{equation}
%where $\bz = [\bg,\by]^T$. The other vectors in Eq.~\eqref{eq:least_squares_red} were defined in Section~\ref{} \footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
%is trivial to distinguish which one is implied from the context. Using the same symbol for these two quantities improves the readability of the paper.}. Note the similarity between Eq.~\eqref{eq:least_squares} and Eq.~\eqref{eq:least_squares_red},
%i.e. redundant calibration can be cast as a non-linear least squares problem.

Analogous to \cite{Smirnov2015}, we now recast Eq.~\eqref{eq:least_squares_red} to 
\begin{equation}
\label{eq:least_squares_complex}
\min_{\breve{\bz}} \|\breve{\br}\| = \min_{\breve{\bz}} \|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
\end{equation}
where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
If we use the Wirtinger derivatives, Eq.~\eqref{eq:wir}, instead of the classic notion of a derivative to define the gradient operator then 
$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole \citep{Smirnov2015}. The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
and complex LM to Eq.~\eqref{eq:least_squares_complex}.

\subsubsection{Complex GN and LM}
\label{sec:complex_GN_LM}
The complex Jacobian is defined to be
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
where 
\begin{align}
\bmJ &= \frac{\partial \bv}{\partial \bz}, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{align}
In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator \citep{Sorber2012}.
The GN update step is now defined as:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
\end{equation}
The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}
where $\bD=\bI\odot\bJ^H\bJ$. We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$.

We use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
parameter vector as follows:
\begin{equation}
\label{eq:update}
\breve{\bz}_{k+1} = \breve{\bz}_{k} + \Delta \breve{\bz}_{k}. 
\end{equation}
We update our parameter vector by using Eq.~\eqref{eq:update} until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
Eq.~\eqref{eq:GN_update} and Eq.~\eqref{eq:LM_update}.

\subsubsection{Analytic Expressions}
\label{sec:analytic}
If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
\begin{equation}
\label{eq:Jacobian_red}
\bJ = \begin{bmatrix}
       \bM & \bN\\
       \conj{\bN} & \conj{\bM}
      \end{bmatrix},
\end{equation}
where
\begin{equation}
\bM =\begin{bmatrix}
      \bO & \bP
     \end{bmatrix},
\end{equation}
and 
\begin{equation}
\bN = \begin{bmatrix}
       \bQ & \bzero
      \end{bmatrix}.
\end{equation}
Moreover,
\begin{equation}
\left [ \bO  \right ]_{\alpha_{pq},j} = \begin{cases}
                                         y_{\phi_{pq}}\conj{g_q} & \textrm{if}~p=j\\
                                         0  & \textrm{otherwise} 
                                        \end{cases},
\end{equation}

\begin{equation}
[\bE]_{ij} = 
\begin{cases}
 \sum_{rs \in \mathcal{RS}_i} \left | g_r \right |^2 \left | g_s \right |^2  & \textrm{if} ~ i=j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
\begin{equation}
[\bF]_{ij} = 
\begin{cases}
 g_i g_j  \left | y_{\zeta_{ij}} \right |^2  & \textrm{if} ~ i \neq j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
[\boldsymbol{G}]_{ij} = 
\begin{cases}
 g_i y_j  \left | g_{\xi_{ij}} \right |^2  & \textrm{if} ~ \xi_{ij}\neq0\\
 0 & \textrm{otherwise}
\end{cases}.
\end{equation}
Moreover, 
\begin{equation}
\mathcal{RS}_i = \left\{rs\in\mathbb{N}^2|(\phi_{rs} = i) \right\}.
\end{equation}

@@

@@

>>>>

\section{Skymodel-based Wirtinger Calibration}
\label{sec:sky_wirtinger}
%In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
%equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
%We present the signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.

%\subsection{Scalar Measurement Equation}
%\label{sec:s_mes}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We estimate the true value of
$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.

@@

Eq.~\eqref{eq:vis_definition} can also be expressed in the following vector form 
\begin{equation}
\label{eq:vis_definition}
\bd = \bv + \bn, 
\end{equation}
where 
\begin{align}
 \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{pq} \conj{g_q},\nonumber\\
 \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}, &  &\label{eq:vec_definitions}
\end{align}
and 
\begin{equation}
\alpha_{pq} =
\begin{cases}
(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
0 & \textrm{otherwise}
\end{cases}.
\end{equation}
The function $\alpha_{pq}$ maps composite antenna indexes to unique single indexes, i.e. 
\begin{equation}
\{\alpha_{12},\alpha_{13},\cdots,\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\}
\end{equation}
and 
\begin{equation}
\{1,2,\cdots,N-1,N,\cdots,B\}
\end{equation}
are equal. We denote the number of baselines in an array with $B$, i.e. $B = \frac{N^2-N}{2}$. The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).

We can now formulate calibration as a least-squares optimization problem:
\begin{equation}
\label{eq:least_squares}
\min_{\bg} \Lambda(\bg) = \min_{\bg} \|\br\|_F^2 = \min_{\bg} \|\bd - \bv(\bg)\|_F^2, 
\end{equation}
where $\Lambda$ is the objective function, $\br$ is the residual vector and $\bg =[g_1,\cdots,g_N]^T$.

The gradient-based minimization algorithms that are generally used to solve least squares problems generally require the model (in Eq.~\eqref{eq:least_squares} the model-predicted vector is denoted by $\bv$) to be differentiable
towards each model parameter (in Eq.~\eqref{eq:least_squares} the model-parameter vector is denoted by $\bg$), i.e. these methods make use of a Jacobian matrix $\bJ$. In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
When the least squares problem is complex it becomes less straightforward to apply these gradient-based minimization methods, as
many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.

The only way to circumvent the differentiability conundrum associated with complex least squares problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown model-parameters separately.

Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
are defined to be 
\begin{align}
\label{eq:wir}
\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ). 
\end{align}
Using the above definitions we can now easily compute the following:
\begin{align}
\frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
\end{align}
If we use the Wirtinger derivative, Eq.~\eqref{eq:wir}, instead of the classic notion of a derivative to define the gradient operator then 
$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole. The Wirtinger derivative can therefore be used to derive complex variants of the real-valued GN and LM algorithms. In these alternative algorithms we treat the complex parameters and their conjugates as separate variables.

\citet{Smirnov2015} recently demonstrated that the Wirtinger derivative can be used to perform normal skymodel-based calibration. We briefly summarize their approach below:
\begin{enumerate}

\item Eq.~\eqref{eq:least_squares} is first recast to
\begin{equation}
\label{eq:least_squares_augmented}
\min_{\breve{\bg}} \Lambda(\breve{\bg}) = \min_{\breve{\bg}} \|\breve{\br}\|_F^2 = \min_{\breve{\bg}} \|\breve{\bd} - \breve{\bv}(\breve{\bg})\|_F^2, 
\end{equation} 
where $\breve{\br} = [\br^T,\conj{\br}^T]^T$, $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$.
\item The complex GN update step is then defined as
\begin{equation}
\label{eq:GN_update_skymodel}
 \Delta \breve{\bg} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br},
\end{equation}
with 
\begin{equation}
\label{eq:Jacobian_skymodel}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
and
\begin{align}
[\bmJ]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial g_i}, & [\bmJ^*]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial \conj{g}_i}. 
\end{align}
Note that the Wirtinger derivative is used in Eq.~\eqref{eq:Jacobian_skymodel} and as such the entries of $\bJ$ are well defined.
\item The complex LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
\begin{equation}
\label{eq:LM_update_skymodel}
\Delta \breve{\bg} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}
where $\bD=\bI\odot\bJ^H\bJ$.
\item Eq.~\eqref{eq:GN_update_skymodel} or Eq.~\eqref{eq:LM_update_skymodel} can now be used iteratively to update the 
parameter vector $\breve{\bg}$:
\begin{equation}
\label{eq:update}
\breve{\bg}_{k+1} = \breve{\bg}_{k} + \Delta \breve{\bg}_{k}. 
\end{equation}
The parameter vector needs to be initialized (an all one vector is one possible choice) and is updated until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
Eq.~\eqref{eq:GN_update_skymodel} and Eq.~\eqref{eq:LM_update_skymodel}.
\end{enumerate}
We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ in this paper. 

@@

\begin{figure*}
\centering
\subfigure[Hexagonal layout]
{\includegraphics[width=0.32\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
\subfigure[Square layout]
{\includegraphics[width=0.32\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
\subfigure[Regular east-west layout]
{\includegraphics[width=0.32\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}

\subfigure[Hexagonal: $\zeta_{pq}$]
{\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
\subfigure[Square: $\zeta_{pq}$]
{\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
\subfigure[Regular east-west: $\zeta_{pq}$]
{\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
\caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
to construct the hexagonal layout, while 100 antennas where used in the square and east-west layouts. In the case of the east-west layout we only plot the positions of the first ten antennas. The maximal amount of redundant baseline groups that can be formed for 
the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.\label{fig:geometry_function}}
\end{figure*}

@@

% \subsection{Limitations}
% \label{sec:scope}
% We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
% is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
% Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}. Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.

As already alluded to in Section~\ref{sec:sky_wirtinger}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately or we can employ the notion of Wirtinger derivatives. The former 
approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.

\subsection{Real and imaginary approach}
\label{sec:ri}
Using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
splitting the problem into its phase-and-amplitude components.

@@

\subsection{Complex Approach}
\label{sec:w}

%\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
%that the Wirtinger derivative can also be applied to redundant calibration.
%We can perform redundant calibration by solving the following optimization problem: 
%\begin{equation}
%\label{eq:least_squares_red}
%\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
%\end{equation}
%where $\bz = [\bg,\by]^T$. The other vectors in Eq.~\eqref{eq:least_squares_red} were defined in Section~\ref{} \footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
%is trivial to distinguish which one is implied from the context. Using the same symbol for these two quantities improves the readability of the paper.}. Note the similarity between Eq.~\eqref{eq:least_squares} and Eq.~\eqref{eq:least_squares_red},
%i.e. redundant calibration can be cast as a non-linear least squares problem.

Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:least_squares_red} to 
\begin{equation}
\label{eq:least_squares_complex}
\min_{\breve{\bz}} \|\breve{\br}\| = \min_{\breve{\bz}} \|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
\end{equation}
where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
and complex LM to Eq.~\eqref{eq:least_squares_complex}.

\subsubsection{Complex GN and LM}
\label{sec:complex_GN_LM}
The complex Jacobian associated with Eq.~\eqref{eq:least_squares_complex} is equal to
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
where 
\begin{equation}
[\bmJ]_{\alpha_{pq},i} = \begin{cases} 
     \frac{\partial v_{pq}}{\partial g_i} & \textrm{if}~i\leq N \\
     \frac{\partial v_{pq}}{\partial y_{i-N}} & \textrm{otherwise}  
\end{cases}, %, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{equation}
and
\begin{equation}
[\bmJ^*]_{\alpha_{pq},i} = \begin{cases} 
     \frac{\partial v_{pq}}{\partial \conj{g}_i} & \textrm{if}~i\leq N \\
     \frac{\partial v_{pq}}{\partial \conj{y}_{i-N}} & \textrm{otherwise}  
\end{cases}. %, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{equation}
The GN update step is now defined as:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
\end{equation}
The LM update rule is defined in a very similar way:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}

As mentioned in Section We use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
parameter vector as follows:
\begin{equation}
\label{eq:update}
\breve{\bz}_{k+1} = \breve{\bz}_{k} + \Delta \breve{\bz}_{k}. 
\end{equation}
We update our parameter vector by using Eq.~\eqref{eq:update} until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
Eq.~\eqref{eq:GN_update} and Eq.~\eqref{eq:LM_update}.

@@

\subsubsection{Signal-to-noise Ratio}
\label{sec:snr}
The following SNR definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

@@

\section{Redundant Wirtinger Calibration - $\bJ$ and $\bH$}
\label{sec:analytic}
If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
\begin{equation}
\label{eq:Jacobian_red}
\bJ = \begin{bmatrix}
       \bM & \bN\\
       \conj{\bN} & \conj{\bM}
      \end{bmatrix},
\end{equation}
where
\begin{equation}
\bM =\begin{bmatrix}
      \bO & \bP
     \end{bmatrix},
\end{equation}
and 
\begin{equation}
\bN = \begin{bmatrix}
       \bQ & \bzero
      \end{bmatrix}.
\end{equation}
Moreover,
\begin{equation}
\left [ \bO  \right ]_{\alpha_{pq},j} = \begin{cases}
                                         y_{\phi_{pq}}\conj{g_q} & \textrm{if}~p=j\\
                                         0  & \textrm{otherwise} 
                                        \end{cases},
\end{equation}

\begin{equation}
[\bE]_{ij} = 
\begin{cases}
 \sum_{rs \in \mathcal{RS}_i} \left | g_r \right |^2 \left | g_s \right |^2  & \textrm{if} ~ i=j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
\begin{equation}
[\bF]_{ij} = 
\begin{cases}
 g_i g_j  \left | y_{\zeta_{ij}} \right |^2  & \textrm{if} ~ i \neq j\\
 0 & \textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
[\boldsymbol{G}]_{ij} = 
\begin{cases}
 g_i y_j  \left | g_{\xi_{ij}} \right |^2  & \textrm{if} ~ \xi_{ij}\neq0\\
 0 & \textrm{otherwise}
\end{cases}.
\end{equation}
Moreover, 
\begin{equation}
\mathcal{RS}_i = \left\{rs\in\mathbb{N}^2|(\phi_{rs} = i) \right\},
\end{equation}
\begin{equation}
\xi_{ij} = 
\begin{cases}
p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
0~\textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
\psi_{ij} = 
\begin{cases}
q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
0~\textrm{otherwise}
\end{cases}.
\end{equation}
